---
title: "Gaudin_ANGSD_Final_Project"
author: "Paul Gaudin"
date: "4/10/2019"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Get Fastq Files

Did this using following SBATCH file: 

```{bash eval=FALSE}

#! /bin/bash -l

#SBATCH --partition=angsd_class
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --job-name=Project_Gets
#SBATCH --time=72:00:00   # HH/MM/SS
#SBATCH --mem=50G       # memory requested, units available: K,M,G,T

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/007/SRR5071107/SRR5071107.fastq.gz /home/ptg4001/Raw_Project_Reads/

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/008/SRR5071108/SRR5071108.fastq.gz /home/ptg4001/Raw_Project_Reads/

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/009/SRR5071109/SRR5071109.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/000/SRR5071110/SRR5071110.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/001/SRR5071111/SRR5071111.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/002/SRR5071112/SRR5071112.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/003/SRR5071113/SRR5071113.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/004/SRR5071114/SRR5071114.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/005/SRR5071115/SRR5071115.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/006/SRR5071116/SRR5071116.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/007/SRR5071117/SRR5071117.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/008/SRR5071118/SRR5071118.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/009/SRR5071119/SRR5071119.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/000/SRR5071120/SRR5071120.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/001/SRR5071121/SRR5071121.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

wget ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR507/002/SRR5071122/SRR5071122.fastq.gz /home/ptg4001/Raw_Project_Reads/ 

```

####Run Fastqc and Multiqc on Fastq files  

SBATCH File: 
```{bash eval=FALSE}
#! /bin/bash -l

#SBATCH --partition=angsd_class
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --job-name=Project_Alignment
#SBATCH --time=72:00:00   # HH/MM/SS
#SBATCH --mem=100G       # memory requested, units available: K,M,G,T

spack load fastqc

fastqc /home/ptg4001/Raw_Project_Reads/*.gz --extract -o /home/ptg4001/Raw_Project_Reads_QC

spack load -r py-multiqc

multiqc -n project_raw.multiqc.html /home/ptg4001/Raw_Project_Reads_QC/.
```

####Run STAR Alignment, Stats, Flagstats, and Multiqc 

Used the following SBATCH File

```{bash eval=FALSE}
#! /bin/bash -l

#SBATCH --partition=angsd_class
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --job-name=Project_Alignment
#SBATCH --time=72:00:00   # HH/MM/SS
#SBATCH --mem=100G       # memory requested, units available: K,M,G,T

spack load star@2.6.1a

spack load samtools@1.9%gcc@6.3.0

for sample in 07 08 09 10 11 12 13 14 15 16 17 18 19 20 21 22; do

	STAR --runMode alignReads --runThreadN 8 --genomeDir /athena/angsd/scratch/Paul_HG_data  --readFilesIn /home/ptg4001/Raw_Project_Reads/SRR50711${sample}.fastq.gz --readFilesCommand zcat --outFileNamePrefix /home/ptg4001/Aligned_Project_Reads/SRR50711${sample}.  --outSAMtype BAM SortedByCoordinate

done

for sample in 07 08 09 10 11 12 13 14 15 16 17 18 19 20 21 22; do

	samtools index /home/ptg4001/Aligned_Project_Reads/SRR50711${sample}.Aligned.sortedByCoord.out.bam

	samtools stats /home/ptg4001/Aligned_Project_Reads/SRR50711${sample}.Aligned.sortedByCoord.out.bam  > /home/ptg4001/Project_Stats/SRR50711${sample}_STAR.stats

	samtools flagstat /home/ptg4001/Aligned_Project_Reads/SRR50711${sample}.Aligned.sortedByCoord.out.bam > /home/ptg4001/Project_Stats/SRR50711${sample}_STAR.flagstats

done

spack load -r py-multiqc

multiqc -n project.multiqc.html /home/ptg4001/Project_Stats/.

```


####Run Feature Counts on Aligned Files

Used the following SBATCH File

```{bash eval=FALSE}

#! /bin/bash -l

#SBATCH --partition=angsd_class
#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --job-name=Project_Alignment
#SBATCH --time=72:00:00   # HH/MM/SS
#SBATCH --mem=100G       # memory requested, units available: K,M,G,T

spack load subread@1.6.2

featureCounts -a /home/ptg4001/hg38.gtf -O -o /home/ptg4001/Project_featurecounts.txt -T 8  /home/ptg4001/Aligned_Project_Reads/*.bam
```

####Import featureCounts readcounts
```{r}
##folder <- "~/Documents/Teaching/ANGSD/RNA/"
folder <- "./"
# reading in featureCounts output
readcounts <- read.table(paste0(folder, "Project_featurecounts.txt"),
header=TRUE)
head(readcounts)

```

#### Just read counts with row names as geneids, rows with sums 0 deleted
```{r}
#library(org.Hs.eg.db)
#library(annotate)

#count_table_adjusted <- readcounts[,7:22]

#row.names(count_table_adjusted) <- getSYMBOL(as.character(readcounts[,1]), data='org.Hs.eg')

#count_table_adjusted = count_table_adjusted[ rowSums(count_table_adjusted)!=0, ] 



```

####Finish preparing count table for analysis in DESeq

```{r}


## retrieve the ENSEMBL symbols for the gene names
#anno.hs <- select(org.Hs.eg.db,
#               keys = gns$V1, 
#               keytype="SYMBOL", 
#               columns=c("ENSEMBL","SYMBOL","GENENAME")) # what to return

gene_trans_ids <- as.character(readcounts[,1])

gene_trans_ids <- gsub("\\..*","", gene_trans_ids)

#AnnotationDbi::select(org.Hs.eg.db, key=gene_trans_ids, columns=c("SYMBOL"), keytype="ENSEMBLPROT")


#library('biomaRt')
#mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
#genes <- as.character(readcounts[,1])
#df<-df[,-4]
#G_list <- getBM(filters= "ensembl_peptide_id", attributes= c("ensembl_peptide_id","hgnc_symbol"),values=genes,mart= mart)
#merge(df,G_list,by.x="gene",by.y="ensembl_peptide_id")

```

```{r}
library(biomaRt)

# 3. Get symbols for Ensembl IDs

ensembl_ids = gene_trans_ids

#ensembl_ids = c("ENSG00000243485", "ENSG00000237613", "ENSG00000186092", "ENSG00000238009")

mart = useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", host = "www.ensembl.org", ensemblRedirect = FALSE)

#mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))

dat = getBM(
  values = ensembl_ids,
  filters = c("ensembl_transcript_id"),
  attributes = c("ensembl_transcript_id", "external_gene_name", "description"),
  mart = mart
)

write.table(
  dat,
  file = "biomart.tsv",
  quote = FALSE,
  sep = "\t",
  row.names = FALSE
)

gene.id.df <- data.frame(dat$external_gene_name, dat$description)

row.names(gene.id.df) <- dat$ensembl_transcript_id

write.csv(gene.id.df, file = "ProjectGenesDataFrame.csv")

```

```{r eval=FALSE}

count_table_adjusted <- readcounts[,7:22]

row_names_as_symbols <- c();

for(i in 1:length(gene_trans_ids)){
  if(is.na(as.character(gene.id.df[gene_trans_ids[i],1]))){
    row_names_as_symbols <- c(row_names_as_symbols, gene_trans_ids[i]);
  }
  else{
    row_names_as_symbols <- c(row_names_as_symbols, (as.character(gene.id.df[gene_trans_ids[i],1])));
  }
}

row.names(count_table_adjusted) <- row_names_as_symbols

count_table_adjusted = count_table_adjusted[ rowSums(count_table_adjusted)!=0, ]

save(row_names_as_symbols, file="row_names_as_symbols.rds")




```

library(org.Hs.eg.db) ## covert gene ids to gene symbols 

  if NA then keep gene ID 

_> one row for each gene 


DESq
